Effectful Semantics for Substructural SSA

### 2023-11-17 pldi24/paper.tex
Another page!

We can now give a semantics for an \isotopessa dialect with types
\(\mc{T}\) and instructions \(\mc{I}\). We'll need:
\begin{itemize}
  \item A \textit{substructural effectful category} \(\mc{C} = \mc{C}_1 \to
  \mc{C}_0\) such that \(\mc{C}_0\) is \textit{distributive} and \textit{Elgot}
  \item A function \(\ms{base}(X)\) from base types \(X \in \mc{T}\) to objects
  in \(|\mc{C}|\), such that \(\ms{a} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Aff}(\mc{C})\) and \(\ms{r} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Rel}(\mc{C})\)
  \item A function \(\ms{inst}(f)\) from instructions \(f \in \mc{I}_p^q(A, B)\)
  to morphisms in \(\mc{C}_p^q(\dnt{A}, \dnt{B})\)
\end{itemize}
We begin by giving a semantics for types and contexts in terms of objects in
\(|\mc{C}|\) in Figure \ref{fig:ssa-type-semantics}: we map tensor products to
tensor products, the unit type to the monoidal unit, and booleans to \(\mb{1} +
\mb{1}\). Contexts are simply interpreted as the tensor products of the variable
types which appear in them; in particular, the quantity annotations have
\textit{no} effect on a context's semantics. We then give a semantics for
structural judgements in Figure \ref{fig:ssa-structural-semantics}. In
particular, we begin by interpreting the judgements that a type is affine (i.e.
\(\tlin{A}{\taff}\)) or relevant (i.e. \(\tlin{A}{\trel}\)) as the drop morphism
\(\ms{drop}(\dnt{A}): \mc{C}(\dnt{A}, I)\) and split morphism
\(\ms{drop}(\dnt{A}): \mc{C}(\dnt{A}, \dnt{A} \otimes \dnt{A})\) respectively.
We can then interpret a context split \(\csplits{\Gamma}{\Delta}{\Xi}\) as a
pure morphism \(\mc{C}^\tint_1(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})\)
which partitions the variables as appropriate, splitting and dropping as
necessary. We interpret a context weakening \(\cwk{\Gamma}{\Delta}\) as the
associated split \(\csplits{\Gamma}{\Delta}{\cdot}\) postcomposed with a unitor.
Finally, we interpret a label-set weakening \(\lwk{\ms{L}}{\ms{K}}\) as a pure
morphism \(\mc{C}^\tint_1(\dnt{\ms{L}}, \dnt{\ms{K}})\), where \rle{join-nil} is
the identity, \rle{join-cons} is the sum of \(\dnt{\lwk{\ms{L}}{\ms{K}}}\) and
\(\dnt{\cwk{\Gamma}{\Delta}}\), and \rle{join-zero} is
\(\dnt{\lwk{\ms{L}}{\ms{K}}}\) postcomposed with the left injection.

@@

\begin{figure}
  \begin{gather*}
    % \boxed{\dnt{\tlin{A}{\taff}: \mc{C}_1^\tint(\dnt{A}, \mb{1})}}
    %   \\
    %   \dnt{\tlin{A}{\taff}} = \ms{drop}(\dnt{A})
    %   % \dnt{\tlin{X}{\taff}} = \ms{drop}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\taff}} 
    %   %   = \dnt{\tlin{A}{\taff}} \otimes \dnt{\tlin{B}{\taff}}
    %   %   ; \lambda 
    %   \\
    % \boxed{\dnt{\tlin{A}{\trel}}
    %   : \mc{C}_1^\tint(\dnt{A}, \dnt{A} \otimes \dnt{A})}
    %   \\
    %   \dnt{\tlin{A}{\trel}} = \ms{split}(\dnt{A})
    %   % \dnt{\tlin{X}{\trel}} = \ms{split}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\trel}} = 
    %   %   \dnt{\tlin{A}{\trel}} \otimes \dnt{\tlin{B}{\trel}}
    %   %   ;\alpha;\dnt{A} \otimes \sigma \otimes \dnt{B};\alpha 
    %   \\
    \boxed{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})}
      \\
    \dnt{\csplits{\cdot}{\cdot}{\cdot}} = \lambda^{-1} \qquad
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta, \thyp{x}{A}{q}}
      {\Xi}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha;\dnt{\Delta} \otimes \sigma;\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} \
      = \dnt{\Gamma} \otimes \ms{split}(\dnt{A});
        \alpha;
        \dnt{\Delta} \otimes \sigma \otimes \dnt{A};
        \alpha 
        \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi}}
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        \otimes \ms{drop}(\dnt{A})
      ; \rho 
      \\
    \boxed{\dnt{\cwk{\Gamma}{\Delta}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta})} 
      \\
      \dnt{\cwk{\Gamma}{\Delta}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\cdot}};\rho 
      \\
    \boxed{\dnt{\lwk{\ms{L}}{\ms{K}}}
      : \mc{C}_1^\tint(\dnt{\ms{L}}, \dnt{\ms{K}})} 
      \\
    \dnt{\lwk{\cdot}{\cdot}} = \ms{id} \qquad
    \dnt{\lwk
      {\ms{L}, \llhyp{\ell}{\Gamma}{A}}
      {\ms{K}, \llhyp{\ell}{\Delta}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}} + \upg{(\dnt{\cwk{\Gamma}{\Delta}} \otimes \dnt{A})}{} 
      \\
    \dnt{\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}};\iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa structural judgements}
  \Description{Semantics for isotope-SSA structural judgements}
  \label{fig:ssa-structural-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\hasty{\Gamma}{p}{a}{A}{q}}
      : \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})} 
      \\
    \dnt{\hasty{\Gamma}{p}{x}{A}{q}} 
      = \dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}
      \qquad
    \dnt{\hasty{\Gamma}{p}{f\;a}{B}{q}}
      = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{p}
      ; \ms{inst}_p(f) 
      \\
    \dnt{\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}};
        \dnt{\hasty{\Delta}{1}{a}{A}{q}} \otimes
        \dnt{\hasty{\Xi}{1}{b}{B}{q}}
      )}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{()}{\mb{1}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}}
      = \upg{(\dnt{\cwk{\Gamma}{\cdot}};\ctt)}{p}
      \qquad
    \dnt{\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}}
      = \upg{(\dnt{\cwk{\Gamma}{\cdot}};\cff)}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{q}}
      )}{p};\dnt{\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      )}{p}
      ; 
      \\ \qquad \qquad \qquad \qquad \qquad \alpha
      ; \dnt{\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
    %   \\
    % \dnt{\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}}
    %   = \dnt{\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}};\alpha_+
  \end{gather*}
  \caption{Semantics for \isotopessa expressions}
  \Description{Semantics for isotope-SSA expressions}
  \label{fig:ssa-term-semantics}
\end{figure}

We can now interpret a term \(\hasty{\Gamma}{p}{a}{A}{q}\) as a
\textit{morphism} taking \(\dnt{\Gamma}\) into \(\dnt{A}\) with
\textit{centrality} \(p\) and \textit{quantity} \(q\) (i.e., lying in
\(\mc{C}_p^q\)). This means that, as expected, a central term becomes a central
morphism in \(\mc{C}_1\), an affine term becomes an affine morphism, and a
relevant term becomes a relevant morphism, and so, in particular, a pure term
becomes a pure morphsim. Our rules for doing so are given in Figure
\ref{fig:ssa-term-semantics}; we will abuse syntax slightly and write
\(\dnt{\text{"judgement"}}\) for \(\dnt{D}\), where \(D\) is a derivation of
"judgement"; in general, it can be proven that the semantics of any two such
derivations are equal. We also define the syntax sugar \(\upg{f}{p}\), where
\(\upg{f}{0}\) is \(\upg{f}{}\) for morphisms in \(\mc{C}_1\) and simply \(f\)
for morphisms in \(\mc{C}_0\), and \(\upg{f}{1}\) is the identity functor on
\(\mc{C}_1\).

Our treatment of \rle{var} is quite standard: the term \(x\) types in \(\Gamma\)
if all other variables in \(\Gamma\) and can be discarded, in which case the
semantics is simply doing so, i.e. \(\dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}\),
lifting to \(\mc{C}_0\) as necessary. Similarly, to type a function application
\rle{var}, we first take \(\dnt{\hasty{\Gamma}{1}{a}{A}{q}}\), which is
guarnateed to be central, lift it to the appropriate category \(\mc{C}_p\), and
then simply postcompose it with the denotation of \(f\). \rle{pair} is handled
by composing the context split \(\csplits{\Gamma}{\Delta}{\Xi}\) taking the
tensor product of the components \(\dnt{\hasty{\Delta}{1}{a}{A}{q}}\),
\(\dnt{\hasty{\Xi}{1}{b}{B}{q}}\); this is well-defined since both components
are central and \(\mc{C}_1\) is monoidal. We then lift the result to the
appropriate \(\mc{C}_p\). The semantics of \rle{unit} is simply dropping the
entire context; \(\dnt{\cwk{\Gamma}{\cdot}}\), while \rle{true}, and \rle{false}
are each handled by first dropping the context and then postcomposing with the
appropriate constant; in all cases, this is followed by lifting to \(\mc{C}_p\).
Finally, \rle{let} and \rle{let2} are handled by first splitting the context,
then passing the right hand component to the semantics of the bound term
\(\dnt{\hasty{\Xi}{1}{a}{A}{q}}\) or \(\dnt{\hasty{\Xi}{1}{a}{A \otimes
B}{q}}\), and then, lifting to \(\mc{C}_p\) as necessary, after reassociating
with \(\alpha\), passing everything to the inner term \(\dnt{\hasty{\Delta,
\thyp{x}{A}{q}}{p}{e}{B}{q}}\) or \(\dnt{\hasty{\Delta, \thyp{x}{A}{q},
\thyp{y}{B}{q}}{p}{e}{C}{q}}\). Note that this inner term is not necessarily
central, unlike the bound term!

@@

Similarly to terms, \textit{blocks} \(\haslb{\Gamma}{t}{\ms{L}}\) are
interpreted as \textit{impure} morphisms from \(\dnt{\Gamma}\) to \(\ms{L}\):
here \(\Gamma\) represents all the variables live on entry to the block, while
\(\ms{L}\) represents all the external blocks which \(t\) might branch to, each
annotated with their live variables and their arguments; our rules are given in
Figure \ref{fig:ssa-block-semantics}. Our semantics for break statements
\(\lbrb{\ell}{a}\) is as follows: we begin by splitting the context into live
variables which pass into dominated blocks and variables used in the argument
via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\). We then pass the latter into the
semantics of the argument (which is pure, and therefore must be lifted to
\(\mc{C}_0\)) \(\dnt{\hasty{\Gamma}{1}{a}{A}{\tint}}\), with the result and live
variables re-associated and passed into
\(\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}\) to obtain the desired target. The
semantics of if-statements are similarly given by first splitting the context
via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\) into variables live in the branches,
\(\Delta\), and variables appearing in the discriminant, \(\Xi\); the latter of
which are passed into the lifted pure morphism
\(\dnt{\hasty{\Delta}{1}{b}{\mb{2}}{\tint}}\). The "if" statement itself is then
simply an application of the distributor \(\delta: A \otimes (\mb{1} + \mb{1})
\to A + A\), with each branch passed to the semantics of its corresponding block
\(\dnt{\haslb{\Xi}{s}{\ms{L}}}\), \(\dnt{\haslb{\Xi}{t}{\ms{L}}}\). The
semantics of \rle{let-blk} and \rle{let2-blk} are analogous to those of
\rle{let} and \rle{let2} in the term case, except that now the bound term \(a\)
is allowed to be impure (and therefore in particular does not need to be
upgraded).

>>>>

We can now give a semantics for an \isotopessa dialect with types
\(\mc{T}\) and instructions \(\mc{I}\). We'll need:
\begin{itemize}
  \item A \textit{substructural effectful category} \(\mc{C} = \mc{C}_1 \to
  \mc{C}_0\) such that \(\mc{C}_0\) is \textit{distributive} and \textit{Elgot}
  \item A function \(\ms{base}(X)\) from base types \(X \in \mc{T}\) to objects
  in \(|\mc{C}|\), such that \(\ms{a} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Aff}(\mc{C})\) and \(\ms{r} \in \ms{lin}(X) \implies \ms{base}(X) \in
  \ms{Rel}(\mc{C})\)
  \item A function \(\ms{inst}(f)\) from instructions \(f \in \mc{I}_p^q(A, B)\)
  to morphisms in \(\mc{C}_p^q(\dnt{A}, \dnt{B})\)
\end{itemize}
We begin by giving a semantics for types and contexts in terms of objects in
Figure \ref{fig:ssa-type-semantics}: we map tensor products to tensor products,
the unit type to the monoidal unit, and booleans to \(\mb{1} + \mb{1}\).
Contexts are simply interpreted as the tensor products of the variable types
which appear in them; in particular, quantity annotations have \textit{no}
effect on a context's semantics. We proceed to give semantics for structural
judgements in Figure \ref{fig:ssa-structural-semantics}. In particular, we
interpret a context split \(\csplits{\Gamma}{\Delta}{\Xi}\) as a pure morphism
\(\mc{C}^\tint_1(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})\) which
partitions the variables as appropriate, splitting and dropping as necessary; a
weakening \(\cwk{\Gamma}{\Delta}\) is interpreted as the associated split
\(\dnt{\csplits{\Gamma}{\Delta}{\cdot}}\) postcomposed with a unitor. Finally, a
label-set weakening \(\lwk{\ms{L}}{\ms{K}}\) is interpreted as a pure morphism
\(\mc{C}^\tint_1(\dnt{\ms{L}}, \dnt{\ms{K}})\), where \rle{join-nil} is the
identity, \rle{join-cons} is the sum of \(\dnt{\lwk{\ms{L}}{\ms{K}}}\) and
\(\dnt{\cwk{\Gamma}{\Delta}}\), and \rle{join-zero} is
\(\dnt{\lwk{\ms{L}}{\ms{K}}}\) postcomposed with the left injection.

@@

\begin{figure}
  \begin{gather*}
    % \boxed{\dnt{\tlin{A}{\taff}: \mc{C}_1^\tint(\dnt{A}, \mb{1})}}
    %   \\
    %   \dnt{\tlin{A}{\taff}} = \ms{drop}(\dnt{A})
    %   % \dnt{\tlin{X}{\taff}} = \ms{drop}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\taff}} 
    %   %   = \dnt{\tlin{A}{\taff}} \otimes \dnt{\tlin{B}{\taff}}
    %   %   ; \lambda 
    %   \\
    % \boxed{\dnt{\tlin{A}{\trel}}
    %   : \mc{C}_1^\tint(\dnt{A}, \dnt{A} \otimes \dnt{A})}
    %   \\
    %   \dnt{\tlin{A}{\trel}} = \ms{split}(\dnt{A})
    %   % \dnt{\tlin{X}{\trel}} = \ms{split}(X) \qquad
    %   % \dnt{\tlin{A \otimes B}{\trel}} = 
    %   %   \dnt{\tlin{A}{\trel}} \otimes \dnt{\tlin{B}{\trel}}
    %   %   ;\alpha;\dnt{A} \otimes \sigma \otimes \dnt{B};\alpha 
    %   \\
    \boxed{\dnt{\csplits{\Gamma}{\Delta}{\Xi}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta} \otimes \dnt{\Xi})}
      \\
    \dnt{\csplits{\cdot}{\cdot}{\cdot}} = \lambda^{-1} \qquad
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta, \thyp{x}{A}{q}}
      {\Xi}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha;\dnt{\Delta} \otimes \sigma;\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}} \otimes \dnt{A};\alpha 
      \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi, \thyp{x}{A}{q}}} \
      = \dnt{\Gamma} \otimes \ms{split}(\dnt{A});
        \alpha;
        \dnt{\Delta} \otimes \sigma \otimes \dnt{A};
        \alpha 
        \\
    \dnt{\csplits
      {\Gamma, \thyp{x}{A}{q}}
      {\Delta}
      {\Xi}}
      = \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        \otimes \ms{drop}(\dnt{A})
      ; \rho 
      \\
    \boxed{\dnt{\cwk{\Gamma}{\Delta}}
      : \mc{C}_1^\tint(\dnt{\Gamma}, \dnt{\Delta})} 
      \qquad \qquad
      \dnt{\cwk{\Gamma}{\Delta}} 
      = \dnt{\csplits{\Gamma}{\Delta}{\cdot}};\rho 
      \\
    \boxed{\dnt{\lwk{\ms{L}}{\ms{K}}}
      : \mc{C}_1^\tint(\dnt{\ms{L}}, \dnt{\ms{K}})} 
      \\
    \dnt{\lwk{\cdot}{\cdot}} = \ms{id} \qquad
    \dnt{\lwk
      {\ms{L}, \llhyp{\ell}{\Gamma}{A}}
      {\ms{K}, \llhyp{\ell}{\Delta}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}} + \upg{(\dnt{\cwk{\Gamma}{\Delta}} \otimes \dnt{A})}{} 
      \\
    \dnt{\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      = \dnt{\lwk{\ms{L}}{\ms{K}}};\iota_0
  \end{gather*}
  \caption{Semantics for \isotopessa structural judgements}
  \Description{Semantics for isotope-SSA structural judgements}
  \label{fig:ssa-structural-semantics}
\end{figure}

\begin{figure}
  \begin{gather*}
    \boxed{\dnt{\hasty{\Gamma}{p}{a}{A}{q}}
      : \mc{C}_p^q(\dnt{\Gamma}, \dnt{A})} 
      \\
    \dnt{\hasty{\Gamma}{p}{x}{A}{q}} 
      = \dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}
      \qquad
    \dnt{\hasty{\Gamma}{p}{f\;a}{B}{q}}
      = \upg{\dnt{\hasty{\Gamma}{1}{a}{A}{q}}}{p}
      ; \ms{inst}_p(f) 
      \\
    \dnt{\hasty{\Gamma}{p}{(a, b)}{A \otimes B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}};
        \dnt{\hasty{\Delta}{1}{a}{A}{q}} \otimes
        \dnt{\hasty{\Xi}{1}{b}{B}{q}}
      )}{p} 
      \\
    \dnt{\hasty{\Gamma}{p}{()}{\mb{1}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p} 
      \quad
    \dnt{\hasty{\Gamma}{p}{\ctt}{\mb{2}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p};\iota_0
      \quad
    \dnt{\hasty{\Gamma}{p}{\cff}{\mb{2}}{q}}
      = \upg{\dnt{\cwk{\Gamma}{\cdot}}}{p};\iota_1 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A}{q}}
      )}{p};\dnt{\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}{q}} 
      \\
    \dnt{\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}{q}}
      = \upg{(
        \dnt{\csplits{\Gamma}{\Delta}{\Xi}}
        ; \dnt{\Delta} \otimes \dnt{\hasty{\Xi}{1}{a}{A \otimes B}{q}}
      )}{p}
      ; 
      \\ \qquad \qquad \qquad \qquad \qquad \alpha
      ; \dnt{\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}{q}}
    %   \\
    % \dnt{\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}{\varnothing}}
    %   = \dnt{\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}};\alpha_+
  \end{gather*}
  \caption{Semantics for \isotopessa expressions}
  \Description{Semantics for isotope-SSA expressions}
  \label{fig:ssa-term-semantics}
\end{figure}

We can now interpret a term \(\hasty{\Gamma}{p}{a}{A}{q}\) as a
\textit{morphism} taking \(\dnt{\Gamma}\) into \(\dnt{A}\) with
\textit{centrality} \(p\) and \textit{quantity} \(q\) (i.e., lying in
\(\mc{C}_p^q\)); our rules for doing so are given in Figure
\ref{fig:ssa-term-semantics}. We will abuse syntax slightly and write
\(\dnt{\text{"judgement"}}\) for \(\dnt{D}\), where \(D\) is a derivation of
"judgement"; in general, it can be proven that the semantics of any two such
derivations are equal. We also define the syntax sugar \(\upg{f}{p}\), where
\(\upg{f}{0}\) is \(\upg{f}{}\) for morphisms in \(\mc{C}_1\) and simply \(f\)
for morphisms in \(\mc{C}_0\), and \(\upg{f}{1}\) is the identity functor on
\(\mc{C}_1\).

Our treatment of \rle{var} is quite standard: semantics of \(x\) is simply
discarding all other variables in \(\Gamma\), i.e.
\(\dnt{\cwk{\Gamma}{\thyp{x}{A}{q}}}\). Similarly, to type a function
application \rle{var}, we first take \(\dnt{\hasty{\Gamma}{1}{a}{A}{q}}\), which
is guaranteed to be central, lift to \(\mc{C}_p\), and then simply postcompose
with \(f\). \rle{pair} is handled by precomposing the context split
\(\csplits{\Gamma}{\Delta}{\Xi}\) with the tensor product of the components
\(\dnt{\hasty{\Delta}{1}{a}{A}{q}}\), \(\dnt{\hasty{\Xi}{1}{b}{B}{q}}\); this is
well-defined since both components are central and \(\mc{C}_1\) is monoidal. The
semantics of \rle{unit} is simply dropping the entire context, while \rle{true}
and \rle{false} are interpreted as first dropping the context and then injecting
into the appropriate branch of the boolean type. Finally, \rle{let} and
\rle{let2} are handled by first splitting the context, then passing the right
hand component to the semantics of the bound term
\(\dnt{\hasty{\Xi}{1}{a}{A}{q}}\) or \(\dnt{\hasty{\Xi}{1}{a}{A \otimes
B}{q}}\), and then, lifting to \(\mc{C}_p\) as necessary, after reassociating
with \(\alpha\), passing everything to the inner term \(\dnt{\hasty{\Delta,
\thyp{x}{A}{q}}{p}{e}{B}{q}}\) or \(\dnt{\hasty{\Delta, \thyp{x}{A}{q},
\thyp{y}{B}{q}}{p}{e}{C}{q}}\). \
% Note that this inner term is not necessarily central, unlike the bound term!

@@

We intrpret \textit{blocks} \(\haslb{\Gamma}{t}{\ms{L}}\) as \textit{impure}
morphisms from \(\dnt{\Gamma}\) to \(\ms{L}\): \(\Gamma\) represents all the
variables live on entry to the block, while \(\ms{L}\) represents all valid
targets, each annotated with live variables and an argument type; our rules are
given in Figure \ref{fig:ssa-block-semantics}. \rle{br} is interpreted by first
splitting the context via \(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\), and then
using \(\Xi\) to compute the target's argument. The result, along the unchanged
live variables \(\Delta\), is then passed to
\(\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}\) to obtain the desired target.
Similarly, we interpret \rle{ite} by first splitting the context via
\(\dnt{\csplits{\Gamma}{\Delta}{\Xi}}\) and using \(\Xi\) to compute the
discriminant. The "if" statement itself is then simply an application of the
distributor \(\delta: \dnt{\Delta} \otimes (\mb{1} + \mb{1}) \to \dnt{\Delta} +
\dnt{\Delta}\); we pass the left of the coproduct to the true branch
\(\dnt{\haslb{\Xi}{s}{\ms{L}}}\) and the right to the false branch
\(\dnt{\haslb{\Xi}{t}{\ms{L}}}\). The semantics of \rle{let-blk} and
\rle{let2-blk} are analogous to those of \rle{let} and \rle{let2}, except that
now the bound term \(a\) is allowed to be impure.
